Modeling trend in temperature volatility using generalized LASSO

### 2018-05-18 nips-2018/paper.tex
add appendix, cut down

% For algorithms
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{hyperref}

@@

\renewcommand*{\figureautorefname}{}%
\renewcommand*{\tableautorefname}{Table}%
\renewcommand*{\sectionautorefname}{Section}%
\renewcommand*{\subsectionautorefname}{Section}%
\renewcommand*{\subsubsectionautorefname}{Section}%

@@

@@

\section{$\ell_1$-trend filtering for estimating variance of a
  time-series}
\label{sec:ell_1-trend-filt}

\attn{Arash: should't the title of this section be "...estimating variance of spatio-temporal data" or something similar?}

$\ell_1$-trend filtering was proposed by \citep{kim_$ell_1$_2009} as a method for estimating a smooth, time-varying trend. It is formulated as the optimization problem
% : a time-series $y_t$, $t=1,...,T$ is observed and we believe that
% it consists of a smooth trend and a stochastic component and the
% goal is to estimate the trend. The estimated trend is desired to be
% smooth, but at the same time, we seek small estimated residuals
% (stochastic component). These two objectives cannot be achieved
% simultaneously and so a trade-off between them should be made. In
% $\ell_1$-trend filtering, this trade-off is formulated as the
% following convex optimization problem: 
\begin{equation}
\min_{\beta} \frac{1}{2} \sum_{t=1}^{T} (y_t-\beta_t)^2+\lambda
\sum_{t=1}^{T-2} \left|\beta_t-2\beta_{t+1}+\beta_{t+2}\right| 
\end{equation}
or equivalently:
\begin{equation}
\min_{\beta} \frac{1}{2} \norm{ y-\beta }_2^2+\lambda \norm{ D\beta}_1
\label{eq:l1tf}
\end{equation}
 where $y_t$ is an observed time-series, $\beta$ is the smooth trend,
 $D$ is a $(T-2)\times T$ matrix, and $\lambda$ is a tuning parameter
 which balances fidelity to the data (small errors in the first term)
 with a desire for smoothness.  
% The \textit{penalty matrix} is:
% \begin{equation}
% D=\begin{bmatrix}
% 1 & -2 & 1 &  &  &  &\\ 
% & 1 & -2 & 1 &  &  &\\ 
% &  & \ddots & \ddots & \ddots  &  &\\ 
% &  & & 1 & -2 & 1 &  \\ 
% &  &  &   & 1 & -2 & 1 
% \end{bmatrix}
% \label{eq:d_matrix}
% \end{equation}
%The first term in \eqref{eq:l1tf} penalizes large residuals while the second term encourages smooth estimated $\beta$.
With the penalty matrix $D$, the estimated $\beta$ will be piecewise
linear. \citep{kim_$ell_1$_2009} proposed a specialized primal-dual
interior point (PDIP) algorithm for solving \eqref{eq:l1tf}. From a
statistical perspective, \eqref{eq:l1tf} is a constrained maximum
likelihood problem with independent observations from a normal
distribution with common variance, $y_t \sim \mbox{N}(\beta_t,
\sigma^2)$, subject to a piecewise linear constraint on $\beta$.

@@

\attn{explain the objective more. give the AR(1) example. Explain what
  you loses by this assumption (ACF,forecasting). Also explain that
  the covariace matrix is diagonal so it cannot capture the covariance
  structure. But incontrast to spatial stat literature, it does not
  make any assumption on estimated variances. Compare to Hallac et al
  and Lingren et al.}

As with~\eqref{eq:l1tf}, one can solve~\eqref{eq:l1tf_var} using the
PDIP algorithm (as in, e.g.,
\texttt{cvxopt}~\citep{andersen_cvxopt:_2013}). In each iteration of PDIP we need to compute a \textit{search direction} by taking a Newton step on a system of nonlinear equations. Appendix B provides details on how this computation can be performed for the objective function \ref{eq:l1tf_var}.

\attn{Arash: I think we should remove Algorithm 1. There are other steps in PDIP which involve adapting the algorithm parameter $w$ and doing Backtracking line search. Also there is only one loop in PDIP.}

@@

The first term in the objective is proportional to the negative
log-likelihood, the second is the temporal penalty for the
time-series at each location $(i,j)$, while the third and fourth,
penalize the difference between the estimated variance of two
vertically and horizontally adjacent points, respectively. The spatial
component of this
penalty is a special case of trend
filtering on graphs~\citep{wang_trend_2016} which penalizes the difference
between the estimated values of the signal on the connected
nodes. As before, we can write
\eqref{eq:l1tf_var_st} in matrix form where $h$ is an $T\times
S$ vector and $D$ is replaced by $D_{TS} \in \mathbb{R}^{(N_t+N_s) \times (T \cdot S)	}$, where $N_t=S \cdot (T-2)$ and $N_s=T \cdot (2n_rn_c-n_r)$ are the number of temporal and spatial constraints, respectively\footnote{$N_s$ is obtained by counting the number of unique constraints at each location and at all times.}. Then, as we
have two different tuning parameters for the temporal and spatial
components, we write $\Lambda =\left[\lambda_1\one_{N_t}^\top,\;
  \lambda_2\one_{N_s}^\top\right]^\top$ leading to \footnote{Throughout the paper, we use $|x|$ for both scalars and vectors. For vectors we use this to denote a vector obtained by taking the absolute value of each entry of $x$.}:
% let $h$ be a
% vector whose first $T$ entries are $h_{11t}$ for $t=1,...,T$, its next
% $T$ entries are $h_{21t}$ and so on, then the optimization problem in
% the matrix form is as
\begin{equation}
\min_h -l(y\given h)+ \Lambda^\top | D_{TS}h |
\label{eq:l1tf_var_st_mat}
\end{equation}
% where $\Lambda = \mathbf{I}_{Tn_rn_c}\left[\lambda_1\one_T^\top,\;
%   \lambda_2\one_{n_rn_c}^\top\right]$ and 
% $D_{total}^t=[D_{t}^t|D_{s}^t]$, where $n_t$ and $n_s$ are the number
% of rows of $D_{t}$ and $D_{s}$, respectively (see below). The matrix
% $D_t$ is the following block-diagonal matrix and corresponds to the
% temporal penalty: 
% \begin{equation}
% D_t=\begin{bmatrix}
% D &  & \\ 
%  & \ddots & \\
%  &  & D
% \end{bmatrix}
% \label{eq:d_t_matrix}
% \end{equation}
%  where $D$ was defined in \eqref{eq:d_matrix}. The number of the
%  diagonal blocks is equal to the grid size $n_r \times n_c$. Each row
%  of the matrix $D_s$ corresponds to one spatial constraint in
%  \eqref{eq:l1tf_var_st}. For example, the first $T$ rows correspond to
%  $|h_{11t}-h_{21t}|$ for $t=1,...,T$, the next $T$ rows correspond to
%  $|h_{11t}-h_{12t}|$, and so on. 
% \attn{Detailed discussion of the PDIP to here}

\attn{Add back the full form of $D$ above? It was too cumbersome
  before, but possibly necessary}

@@

\attn{We note that stochastic gradient descent could be used but...}

\attn{Arash: The primal objective is not differentiable and so we cannot use SGD. The dual has constriants and so the plane SGD cannot be used. We can use projected SGD but I think it would converge slowly. Since other papers on LASSO have not considered SGD I don't think we need to justify why we did not use it.}

@@

Consider an optimization problem of the form $\min_h f(h)$, where $h\in\mathbb{R}^n$ is the \textit{global variable} and $f(h):\mathbb{R}^n \rightarrow \mathbb{R}\cup \{+\infty\}$ is convex. The goal in consensus optimization is to break this problem into several smaller sub-problems which can be solved independently in each iteration of optimization.

Assume that it is possible to define a set of \textit{local variables} $x_i \in \mathbb{R}^{n_i}$ such that $f(h)=\sum_i f_i(x_i)$, where each $x_i$ is a subset of the global variable $h$. More specifically, each entry of the local variables corresponds to an entry of the global variable. Therefore we can define a mapping $\mathscr{G}(i,j)$ from the local variables indices into the global variable indices: $k=\mathscr{G}(i,j)$ means that the $j^\text{th}$ entry of $x_i$ is $h_k$ (or $(x_i)_j=h_k$). For ease of notation, define $\tilde{h}_i \in \mathbb{R}^{n_i}$ as $(\tilde{h}_i)_j=h_{\mathscr{G}(i,j)}$. Then, the original optimization problem is equivalent to the following problem:

\begin{equation}
\begin{aligned}
\min_{\{x_1,...,x_N  \}} \sum_i f_i(x_i)\\
 s.t. \,\,\, \tilde{h}_i=x_i \,\,\,\,\,\,
\end{aligned}
\label{eq:consADMM}
\end{equation}

It is important to note that each entry of the global variable may correspond to several entries of the local variables and so the constraints $\tilde{h}_i=x_i$ enforce the consensus between the local variables corresponding to the same global variable.

The \textit{augmented Lagrangian} corresponding to the problem \ref{eq:consADMM} is $L_\rho(x,h,y)=\sum_i \big(f_i(x_i)+u_i^\top(x_i-\tilde{h}_i) + (\rho/2) \lVert x_i-\tilde{h}_i \lVert_2^2 \big)$. Now, we can apply ADMM to $L_\rho$ which results in the following ADMM updating steps at each iteration $m$:

\begin{equation}
\begin{aligned}
& x_i^{m+1}:=\argmin_{x_i} \bigg( f_i(x_i) + (u_i^m)^\top x_i + (\rho/2) \lVert x_i-\tilde{h}_i^m \lVert_2^2\bigg)\\
& h_k^{m+1}:=(1/S_k)\sum_{\mathscr{G}(i,j)=k} (x_i^{m+1})_j\\
& u_i^{m+1}:=u_i^m + \rho (x_i^{m+1}-\tilde{h}_i^{m+1})
\end{aligned}
\label{eq:ADMM_steps}
\end{equation}

Here, $S_k$ is the number of local variable entries that correspond to
$h_k$, and $u_i$ are the Lagrange multipliers.

\begin{comment}
The $x$-update is
often a simple modification as the original optimization, though over only a small
partition, the $h$-update simply averages over collections of
analogous local variables while the $u$ update enforces sharing. Both
the $x$ and $u$ updates are parallelizable over the partition of local
variables.
\end{comment}

\attn{I don't much like this explanation or notation. Seems unclear}

\attn{Arash: I re-wrote this section.}

To solve the optimization problem \eqref{eq:l1tf_var_st_mat} or
\eqref{eq:l1tf_var_st} using this method, we need to address two
questions: first, how to choose the 
local variables $x_i$, and second, how to solve the optimization
problem for updating these variables (the first line of
\eqref{eq:ADMM_steps} which we will refer to it as \textit{x-update step}).

In \autoref{fig:data_cube}, the global variable $h$ is represented
as a cube (using the subset of the US as an example). We decompose $h$
into sub-cubes as shown in the figure by white lines. It is easy to see
that by this definition of $x_i$, the objective
\eqref{eq:l1tf_var_st_mat} decomposes as $\sum_i f_i(x_i)$ where
$f_i(x_i)=-l(y_i\given x_i)+\Lambda_{(i)}^\top |D_{(i)}x_i|$, and
$\Lambda_{(i)}$ and $D_{(i)}$ contain the temporal and spatial
penalties corresponding to $x_i$ only. The x-update step in \autoref{eq:ADMM_steps} is the following optimization problem: $x_i^{m+1}:=\argmin_{x_i} ( f_i(x_i) +
(u_i^m)^\top x_i + (\rho/2) \lVert x_i-\tilde{z}_i^m \lVert_2^2)$.

\attn{I would like to be able to write all this in terms of $h$ instead of $x$}

\attn{Arash: now I'm using $h$ as the global variables. We cannot write the local variables in terms of $h$ since they are different from $h$.}

\begin{figure}[tb]
  \centering
  \includegraphics[height=.2\textheight]{Figures/data_cube}
    \caption{The cube represents the global variable $h$ in space and
      time. The sub-cubes specified by the white lines are
      $x_i$. \attn{Can we use $h_i$ here instead of $x_i$?}}
    \label{fig:data_cube}
\end{figure}

By this definition of $x_i$, the update step for $x_i$ is the
following optimization problem: $x_i^{m+1}:=\argmin_{x_i} ( f_i(x_i) +
(u_i^m)^\top x_i + (\rho/2) \lVert x_i-\tilde{z}_i^m \lVert_2^2)$. We solve this using the PDIP method. As it was explained in \autoref{sec:l1tf_var}, to use PDIP we need to compute the dual problem, which in turn needs the computation of the Fenchel conjugate of the loss function. In addition, referring to \autoref{alg:pdip}, we need to compute the gradient and Jacobian of the conjugate functions. It is important to note that compare to the optimization problem \autoref{eq:l1tf_var}, the loss function in the x-update step in \autoref{eq:ADMM_steps} includes the quadratic term $(\rho/2) \lVert x_i-\tilde{z}_i^m \lVert_2^2$. This makes the computation more involved than \autoref{sec:l1tf_var}. The details of the computations are explained in \autoref{sec:app_consADMM}.\attn{change this to appendix}

\attn{I'm not sure if we want to devote some space to write this as a separate algorithm. The steps are those mentioned in \ref{eq:ADMM_steps} and the x-update is now explained in the appendix. So we may want to remove algorithm \ref{alg:ADMM}. }

The complete ADMM algorithm for estimating the variances is
represented in \autoref{alg:ADMM}. All the computations in the three
updating steps \eqref{eq:ADMM_steps} can be performed in parallel. The
number of rows and columns of the sub-cubes should be chosen so that
the updating of $x_i$ could be performed in one processor. We choose
$3\times3\times521$ sub-cubes.

\begin{algorithm}[tb]
  \caption{ADMM for sparse estimation of variance of spatio-temporal
    data \attn{Fix this}}
  \label{alg:ADMM}
  \begin{algorithmic}
    \STATE {\bfseries Input:} data $y$, mapping $\mathscr{G}(i,j)$,
    $\rho,\lambda_t,\lambda_s$ 
    \STATE Initialization: $x_i^0=z^0=u_i^0=\textbf{0}$.
    \FOR{$m=1,2,...$ }
    \FOR{$i=1$ {\bfseries to} $N_{sub-cubes}$}
    \STATE compute $\nu_i$ from \eqref{eq:dual_x_update}
    \STATE compute $w_i$ from \eqref{eq:conj_x_update}
    \STATE set $x_i^m:=w_i$
    \ENDFOR
    \STATE Compute $z^m$ from  \eqref{eq:ADMM_steps}
    \STATE Compute $u_i^m$ from  \eqref{eq:ADMM_steps}
    \ENDFOR
  \end{algorithmic}
\end{algorithm}

Because \autoref{alg:ADMM} breaks the large optimization into
sub-problems that can be solved independently, it is amenable to a
split-gather parallelization strategy via, e.g., the map reduce framework.
In each iteration, the
computation time will be equal to the time to solve each sub-problem
plus the time to communicate the solutions on the master processor
and perform the consensus step. Since each sub-problem is
small, with parallelization, the computation time in each iteration
will be small. In addition, our experiments with several values of
$\lambda_t$ and $\lambda_s$ showed that the algorithm converges in few
hundreds iterations. 
\attn{Need to redo this:}Solving each sub-problem on a machine with four
3.20GHz Intel i5-3470 cores takes less than 3 seconds on average, and
so for example if we assume that communication time is 10 seconds and
the algorithm converges in 300 iterations, with parallelization on
$N_{sub-cubes}$ machines, the algorithm will converge in about 1
hour. Assuming that we use $N_{sub-cubes}$ machines and that the
convergence rate of the algorithm is independent of the grid size,
this time will be independent of the grid size.

If we perform these computations on a single machine, the computation
time grows linearly with $N_{sub-cubes}$. For example, for the data in
a grid over the united states and using $3\times3\times521$ sub-cubes
each iteration of the algorithm will take about 20 minutes on a single
machine and so with 300 iterations it will take several days to
converge. Given that we need to compute the solution for several
values of the parameters $\lambda_t$ and $\lambda_s$, this computation
time is not feasible.

Therefore, this algorithm is only useful if we can parallelize the
computation over several machines. In the next section, we describe
another algorithm which makes the computation feasible on a single
machine.

@@

\attn{Need a good dummy variable here, it can't be $u$.} Consider the generic optimization problem
$\min_x f(x)+g(Dx)$
where $x\in \mathbb{R}^n$ and $D\in \mathbb{R}^{m\times n}$. Each
iteration of the linearized ADMM
algorithm~\citep{parikh_proximal_2014} for solving this problem 
has the form

\begin{align}
x & \leftarrow \prox_{\mu f} \left(x - (\mu/\rho)D^\top (D x - z + u )\right)\\
z & \leftarrow \prox_{\rho g} \left(z + u\right) \label{eq:linADMM_steps}\\
u & \leftarrow u + D x - z
\end{align}

\noindent where the algorithm parameters $\mu$ and $\rho$ satisfy $0 < \mu < \rho/\norm{D}_2^2$, $z,u\in \mathbb{R}^m$ and the proximal operator is defined as

\begin{equation}
 \prox_{\alpha f}(u) = \min_x \,\, \alpha \cdot f(x)+\frac{1}{2} \norm{ x-u}_2^2.
\label{eq:linADMM_prox}
\end{equation}

\attn{What are $\mu$ and $\rho$?}

\attn{Arash: I explained it above. They are algorithm parameters.}

Clearly, \eqref{eq:l1tf_var_st} has this form necessary for using this algorithm.
% The optimization problem 
% form \eqref{eq:linADMM_opt} by defining
% \begin{align}
% & f(x):= \sum_{k} f_k(x_k) := \sum_{k}x_k+y_{k}^2e^{-x_{k}}\\
% & g(z):= \sum_{l} g_l(z_l) := \sum_{l} \lambda_l |z_l|\label{eq:linADMM_fg}\\
% & z = Dx
% \end{align}
% where $y_k$ is the $k^{th}$ entry of the vector whose entries are
% $y_{ijt}$, and $\lambda_l$ is the $l^{th}$ entry of the vector
% $\Lambda^\top=(\lambda_t\textbf{e}_{n_t}^\top|\lambda_s\textbf{e}_{n_s}^\top)$
% (see \autoref{sec:exten}). 
To perform the steps in \eqref{eq:linADMM_steps}, we need to evaluate
$\prox_{\mu f}$ and $\prox_{\rho g}$. Proximal
algorithms are feasible only if these proximal operators can be
evaluated efficiently which, as we show next, is the case for our
problem.

>>>>

% For algorithms
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{hyperref}
\hypersetup{backref,colorlinks=true,citecolor=blue,linkcolor=blue,urlcolor=blue}

@@

\renewcommand*{\figureautorefname}{Figure}%
\renewcommand*{\tableautorefname}{Table}%
\renewcommand*{\sectionautorefname}{Section}%
\renewcommand*{\subsectionautorefname}{Section}%
\renewcommand*{\subsubsectionautorefname}{Section}%

@@

\usepackage{xr}
\externaldocument{paper-supplement}

@@

\section{Estimating the variance of spatio-temporal data}
\label{sec:ell_1-trend-filt}

$\ell_1$-trend filtering was proposed by \citep{KimKoh2009} as a method for estimating a smooth, time-varying trend. It is formulated as the optimization problem
% : a time-series $y_t$, $t=1,...,T$ is observed and we believe that
% it consists of a smooth trend and a stochastic component and the
% goal is to estimate the trend. The estimated trend is desired to be
% smooth, but at the same time, we seek small estimated residuals
% (stochastic component). These two objectives cannot be achieved
% simultaneously and so a trade-off between them should be made. In
% $\ell_1$-trend filtering, this trade-off is formulated as the
% following convex optimization problem: 
\begin{equation}
\min_{\beta} \frac{1}{2} \sum_{t=1}^{T} (y_t-\beta_t)^2+\lambda
\sum_{t=1}^{T-2} \left|\beta_t-2\beta_{t+1}+\beta_{t+2}\right| 
\end{equation}
or equivalently:
\begin{equation}
\min_{\beta} \frac{1}{2} \norm{ y-\beta }_2^2+\lambda \norm{ D\beta}_1
\label{eq:l1tf}
\end{equation}
 where $y_t$ is an observed time-series, $\beta$ is the smooth trend,
 $D$ is a $(T-2)\times T$ matrix, and $\lambda$ is a tuning parameter
 which balances fidelity to the data (small errors in the first term)
 with a desire for smoothness.  
% The \textit{penalty matrix} is:
% \begin{equation}
% D=\begin{bmatrix}
% 1 & -2 & 1 &  &  &  &\\ 
% & 1 & -2 & 1 &  &  &\\ 
% &  & \ddots & \ddots & \ddots  &  &\\ 
% &  & & 1 & -2 & 1 &  \\ 
% &  &  &   & 1 & -2 & 1 
% \end{bmatrix}
% \label{eq:d_matrix}
% \end{equation}
%The first term in \eqref{eq:l1tf} penalizes large residuals while the second term encourages smooth estimated $\beta$.
With the penalty matrix $D$, the estimated $\beta$ will be piecewise
linear. \citep{KimKoh2009} proposed a specialized primal-dual
interior point (PDIP) algorithm for solving \eqref{eq:l1tf}. From a
statistical perspective, \eqref{eq:l1tf} is a constrained maximum
likelihood problem with independent observations from a normal
distribution with common variance, $y_t \sim \mbox{N}(\beta_t,
\sigma^2)$, subject to a piecewise linear constraint on $\beta$.

@@

As with~\eqref{eq:l1tf}, one can solve~\eqref{eq:l1tf_var} using the
PDIP algorithm (as in, e.g.,
\texttt{cvxopt}~\citep{andersen_cvxopt:_2013}). In each iteration of
PDIP we need to compute a search direction by taking a Newton
step on a system of nonlinear equations. Due to space limitations, we
defer details to Appendix B in the Supplement.

@@

The first term in the objective is proportional to the negative
log-likelihood, the second is the temporal penalty for the
time-series at each location $(i,j)$, while the third and fourth,
penalize the difference between the estimated variance of two
vertically and horizontally adjacent points, respectively. The spatial
component of this
penalty is a special case of trend
filtering on graphs~\citep{WangSharpnack2016} which penalizes the difference
between the estimated values of the signal on the connected
nodes. As before, we can write
\eqref{eq:l1tf_var_st} in matrix form where $h$ is an $T\times
S$ vector and $D$ is replaced by $D_{TS} \in \mathbb{R}^{(N_t+N_s)
  \times (T \cdot S)	}$, where $N_t=S \cdot (T-2)$ and $N_s=T \cdot
(2n_rn_c-n_r)$ are the number of temporal and spatial constraints,
respectively\footnote{$N_s$ is obtained by counting the number of
  unique constraints at each location and at all times.}. Then, as we 
have two different tuning parameters for the temporal and spatial
components, we write $\Lambda =\left[\lambda_1\one_{N_t}^\top,\;
  \lambda_2\one_{N_s}^\top\right]^\top$ leading
to:\footnote{Throughout the paper, we use $|x|$ for both scalars and
  vectors. For vectors we use this to denote a vector obtained by
  taking the absolute value of each entry of $x$.} 
% let $h$ be a
% vector whose first $T$ entries are $h_{11t}$ for $t=1,...,T$, its next
% $T$ entries are $h_{21t}$ and so on, then the optimization problem in
% the matrix form is as
\begin{equation}
\min_h -l(y\given h)+ \Lambda^\top | D_{TS}h |.
\label{eq:l1tf_var_st_mat}
\end{equation}
\attn{equations should have proper punctuation as if they were a sentence.}
% where $\Lambda = \mathbf{I}_{Tn_rn_c}\left[\lambda_1\one_T^\top,\;
%   \lambda_2\one_{n_rn_c}^\top\right]$ and 
% $D_{total}^t=[D_{t}^t|D_{s}^t]$, where $n_t$ and $n_s$ are the number
% of rows of $D_{t}$ and $D_{s}$, respectively (see below). The matrix
% $D_t$ is the following block-diagonal matrix and corresponds to the
% temporal penalty: 
% \begin{equation}
% D_t=\begin{bmatrix}
% D &  & \\ 
%  & \ddots & \\
%  &  & D
% \end{bmatrix}
% \label{eq:d_t_matrix}
% \end{equation}
%  where $D$ was defined in \eqref{eq:d_matrix}. The number of the
%  diagonal blocks is equal to the grid size $n_r \times n_c$. Each row
%  of the matrix $D_s$ corresponds to one spatial constraint in
%  \eqref{eq:l1tf_var_st}. For example, the first $T$ rows correspond to
%  $|h_{11t}-h_{21t}|$ for $t=1,...,T$, the next $T$ rows correspond to
%  $|h_{11t}-h_{12t}|$, and so on. 
% \attn{Detailed discussion of the PDIP to here}

% \attn{Add back the full form of $D$ above? It was too cumbersome
%   before, but possibly necessary}

@@

@@

Given an optimization problem of the form $\min_h f(h)$, where
$h\in\mathbb{R}^n$ is the global variable and
$f(h):\mathbb{R}^n \rightarrow \mathbb{R}\cup \{+\infty\}$ is
convex. Consensus optimization breaks this problem
into several smaller sub-problems that can be solved independently in
each iteration of optimization.

Assume it is possible to define a set of local variables
$x_i \in \mathbb{R}^{n_i}$ such that $f(h)=\sum_i f_i(x_i)$, where
each $x_i$ is a subset of the global variable $h$. More specifically,
each entry of the local variables corresponds to an entry of the
global variable. Therefore we can define a mapping $\mathscr{G}(i,j)$
from the local variable indices into the global variable indices:
$k=\mathscr{G}(i,j)$ means that the $j^\text{th}$ entry of $x_i$ is
$h_k$ (or $(x_i)_j=h_k$). For ease of notation, define $\tilde{h}_i
\in \mathbb{R}^{n_i}$ as $(\tilde{h}_i)_j=h_{\mathscr{G}(i,j)}$. Then,
the original optimization problem is equivalent to the following
problem:  
\begin{equation}
\begin{aligned}
\min_{\{x_1,...,x_N  \}} \sum_i f_i(x_i)\\
 s.t. \,\,\, \tilde{h}_i=x_i. \,\,\,\,\,\,
\end{aligned}
\label{eq:consADMM}
\end{equation}
It is important to note that each entry of the global variable may
correspond to several entries of the local variables and so the
constraints $\tilde{h}_i=x_i$ enforce the consensus between the local
variables corresponding to the same global variable.

The augmented Lagrangian corresponding to 
\eqref{eq:consADMM} is $L_\rho(x,h,y)=\sum_i
\big(f_i(x_i)+u_i^\top(x_i-\tilde{h}_i) + (\rho/2) \lVert
x_i-\tilde{h}_i \lVert_2^2 \big)$. Now, we can apply ADMM to $L_\rho$
which results in the following ADMM updates:
\begin{align}
x_i&\leftarrow\argmin_{x_i} f_i(x_i) + (u_i)^\top x_i +
     (\rho/2)  \norm{ x_i-\tilde{h}_i }_2^2\\
h_k&\leftarrow (1/S_k)\sum_{\mathscr{G}(i,j)=k} (x_i)_j \label{eq:ADMM_steps}\\
u_i&\leftarrow u_i + \rho (x_i-\tilde{h}_i).
\end{align}
\attn{If you hate the left arrow, that's fine. But change it
  everywhere, and use $=$. The $:=$ notation means that ``thing on
  left is defined to be thing on right'' and isn't really appropriate here.}
Here, $S_k$ is the number of local variable entries that correspond to
$h_k$, and $u_i$ are the Lagrange multipliers.

To solve the optimization problem \eqref{eq:l1tf_var_st_mat} using
this method, we need to address two 
questions: first, how to choose the 
local variables $x_i$, and second, how to the update in the first line of
\eqref{eq:ADMM_steps}.

\begin{figure}[tb]
  \centering
  \includegraphics[height=.2\textheight]{Figures/data_cube}
  \caption{The cube represents the global variable $h$ in space and
    time. The sub-cubes specified by the white lines are
    $x_i$.}
  \label{fig:data_cube}
\end{figure}

In \autoref{fig:data_cube}, the global variable $h$ is represented
as a cube (using the subset of the US as an example). We decompose $h$
into sub-cubes as shown by white lines. With
this definition of $x_i$, the objective
\eqref{eq:l1tf_var_st_mat} decomposes as $\sum_i f_i(x_i)$ where
$f_i(x_i)=-l(y_i\given x_i)+\Lambda_{(i)}^\top |D_{(i)}x_i|$, and
$\Lambda_{(i)}$ and $D_{(i)}$ contain the temporal and spatial
penalties corresponding to $x_i$ only. 
% The $x$-update step in
% \eqref{eq:ADMM_steps} is the following optimization problem:
% $x_i^{m+1}:=\argmin_{x_i} ( f_i(x_i) + 
% (u_i^m)^\top x_i + (\rho/2) \lVert x_i-\tilde{z}_i^m \lVert_2^2)$.
Thus, with this choice of the local variables $x_i$, we solve the
$x$-update using the PDIP method. 
% As it 
% was explained in \autoref{sec:l1tf_var}, to use PDIP we need to
% compute the dual problem, which in turn needs the computation of the
% Fenchel conjugate of the loss function. In addition, referring to
% \autoref{alg:pdip}, we need to compute the gradient and Jacobian of
% the conjugate functions. It is important to note that compare to the
% optimization problem \autoref{eq:l1tf_var}, the loss function in the
% x-update step in \autoref{eq:ADMM_steps} includes the quadratic term
% $(\rho/2) \lVert x_i-\tilde{z}_i^m \lVert_2^2$. This makes the
% computation more involved than \autoref{sec:l1tf_var}. 
The details of the remaining computations
the computations are explained in the Supplement.

% The complete ADMM algorithm for estimating the variances is
% represented in \autoref{alg:ADMM}. All the computations in the three
% updating steps \eqref{eq:ADMM_steps} can be performed in parallel. The
% number of rows and columns of the sub-cubes should be chosen so that
% the updating of $x_i$ could be performed in one processor. We choose
% $3\times3\times521$ sub-cubes.

% \begin{algorithm}[tb]
%   \caption{ADMM for sparse estimation of variance of spatio-temporal
%     data \attn{Fix this}}
%   \label{alg:ADMM}
%   \begin{algorithmic}
%     \STATE {\bfseries Input:} data $y$, mapping $\mathscr{G}(i,j)$,
%     $\rho,\lambda_t,\lambda_s$ 
%     \STATE Initialization: $x_i^0=z^0=u_i^0=\textbf{0}$.
%     \FOR{$m=1,2,...$ }
%     \FOR{$i=1$ {\bfseries to} $N_{sub-cubes}$}
%     \STATE compute $\nu_i$ from \eqref{eq:dual_x_update}
%     \STATE compute $w_i$ from \eqref{eq:conj_x_update}
%     \STATE set $x_i^m:=w_i$
%     \ENDFOR
%     \STATE Compute $z^m$ from  \eqref{eq:ADMM_steps}
%     \STATE Compute $u_i^m$ from  \eqref{eq:ADMM_steps}
%     \ENDFOR
%   \end{algorithmic}
% \end{algorithm}

Because consensus ADMM breaks the large optimization into
sub-problems that can be solved independently, it is amenable to a
split-gather parallelization strategy via, e.g., the map reduce framework.
In each iteration, the
computation time will be equal to the time to solve each sub-problem
plus the time to communicate the solutions on the master processor
and perform the consensus step. Since each sub-problem is
small, with parallelization, the computation time in each iteration
will be small. In addition, our experiments with several values of
$\lambda_t$ and $\lambda_s$ showed that the algorithm converges in few
hundreds iterations. 
% \attn{Need to redo this:}Solving each sub-problem on a machine with four
% 3.20GHz Intel i5-3470 cores takes less than 3 seconds on average, and
% so for example if we assume that communication time is 10 seconds and
% the algorithm converges in 300 iterations, with parallelization on
% $N_{sub-cubes}$ machines, the algorithm will converge in about 1
% hour. Assuming that we use $N_{sub-cubes}$ machines and that the
% convergence rate of the algorithm is independent of the grid size,
% this time will be independent of the grid size. 
% If we perform these computations on a single machine, the computation
% time grows linearly with $N_{sub-cubes}$. For example, for the data in
% a grid over the united states and using $3\times3\times521$ sub-cubes
% each iteration of the algorithm will take about 20 minutes on a single
% machine and so with 300 iterations it will take several days to
% converge. Given that we need to compute the solution for several
% values of the parameters $\lambda_t$ and $\lambda_s$, this computation
% time is not feasible. 
However, this algorithm is only useful if we can parallelize the
computation over several machines. In the next section, we describe
another algorithm which makes the computation feasible on a single
machine.

@@

Consider the generic optimization problem
$\min_x f(x)+g(Dx)$
where $x\in \mathbb{R}^n$ and $D\in \mathbb{R}^{m\times n}$. Each
iteration of the linearized ADMM
algorithm~\citep{parikh_proximal_2014} for solving this problem 
has the form
\begin{align}
x & \leftarrow \prox_{\mu f} \left(x - (\mu/\rho)D^\top (D x - z + u )\right)\\
z & \leftarrow \prox_{\rho g} \left(z + u\right) \label{eq:linADMM_steps}\\
u & \leftarrow u + D x - z
\end{align}
where the algorithm parameters $\mu$ and $\rho$ satisfy $0 < \mu < \rho/\norm{D}_2^2$, $z,u\in \mathbb{R}^m$ and the proximal operator is defined as
\begin{equation}
 \prox_{\alpha f}(u) = \min_x \,\, \alpha \cdot f(x)+\frac{1}{2} \norm{ x-u}_2^2.
\label{eq:linADMM_prox}
\end{equation}

%Clearly, \eqref{eq:l1tf_var_st} has this form necessary for using this algorithm.
% The optimization problem 
% form \eqref{eq:linADMM_opt} by defining
% \begin{align}
% & f(x):= \sum_{k} f_k(x_k) := \sum_{k}x_k+y_{k}^2e^{-x_{k}}\\
% & g(z):= \sum_{l} g_l(z_l) := \sum_{l} \lambda_l |z_l|\label{eq:linADMM_fg}\\
% & z = Dx
% \end{align}
% where $y_k$ is the $k^{th}$ entry of the vector whose entries are
% $y_{ijt}$, and $\lambda_l$ is the $l^{th}$ entry of the vector
% $\Lambda^\top=(\lambda_t\textbf{e}_{n_t}^\top|\lambda_s\textbf{e}_{n_s}^\top)$
% (see \autoref{sec:exten}). 
% To perform the steps in \eqref{eq:linADMM_steps}, we need to evaluate
% $\prox_{\mu f}$ and $\prox_{\rho g}$. 
Proximal
algorithms are feasible when these proximal operators can be
evaluated efficiently which, as we show next, is the case for our
problem.
